\section{Appendix}
\subsection{Math Glossary for lowly CS majors}
\begin{defn}[Orthogonal Projection Matrix]
A matrix $\A \in \R^{n \times n}$ is an orthogonal projection matrix if $\A^2 = \A$ and $\A^\top = \A$.
\end{defn}
\begin{defn}[Gamma Function]
The Gamma Function $\Gamma(x): \R \to \R$ is defined as
\begin{align*}
    \Gamma(x) = \int_0^\infty t^{x-1}e^{-t} {\rm d}t
\end{align*}
Importantly, it satisfies $\Gamma(x) = (x-1)!$ for $x \in \Z^+$. In general, it satisfies $\Gamma(x) = x \cdot \Gamma(x-1)$.
\end{defn}
\begin{defn}[Hyperplane]
In $d$ dimensions, a hyperplane is a subspace of dimension $d-1$.
\end{defn}
\begin{defn}[Half-space]
A half space is either of the two parts that a hyperplane divides a space into.
\end{defn}
\begin{example}[half-space]
In $2$ dimensions, consider any line in the plane. Then the parts on either side of the line define half-spaces.
\end{example}
\begin{defn}[Orthant]
An orthant in $d$ dimensions is defined to be the intersection of $d$ mutually orthogonal half-spaces. The halfspaces induced by the standard bases for $\R^d$ induce $2^d$ orthants (for $d=2$, quadrants).
\end{defn}
\begin{example}[Orthants]
The upper right quadrant in $2$ dimensions is the intersection of the half-spaces $\set{(x,y) \in \R^2 \mid x \geq 0 }$ and $\set{(x,y) \in \R^2 \mid y \geq 0}$. 
\end{example}
\begin{defn}[Smoothness]
A function $f: A \to B$ is said to be smooth if it is infinitely-differentiable everywhere.
\end{defn}
\begin{defn}[Latent]
A variable which is inferred, but not observed is considered to be a latent variable.
\end{defn}
\begin{example}[Latent Variable]
Properties such as time taken to add to cart, time taken to check out cart, hit rate of recommendations etc are all measurable, but ``consumer satisfaction'' scores are inferred from these, not observed.
\end{example}
\subsection{Classical Interpolation Methods}
\label{sec:app:interp}
\subsubsection{Linear Interpolation}
\subsubsection{Polynomial Interpolation}
\subsubsection{Spine Interpolation}

\subsection{Lagrange Multipliers}
